---
title: 'Lab Week 5: Time Series and Map'
author: "Jessica Jagdeo"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Attach packages

```{r}

library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)

```

## Monthly US energy consumption (renewables)

```{r}

us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
  

```

- Make description all lowercase
- Only keep observations where "description" variable contains "consumption"
- Remove any observations where "description" variable contains "total"

```{r}

renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total")) # Removes characters with "total" in the name

```

### Convert 'yyyymm' column to a date

```{r}

renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>%  # Separate day from month and year
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# Make a version where I have the month & year in separate columns

renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))

```

### Look at it:

```{r}

renew_gg <- ggplot(renew_date, aes(x = month_sep, 
                                   y = value,
                                   group = description)) +
  geom_line(aes(color = description)) 
  

renew_gg

```

Updating colors with paletteer paletters:
```{r}

renew_gg + 
  scale_color_paletteer_d("LaCroixColoR::Coconut") # Choosing "d" for discrete

```

### Coerce renew_parsed to a tsibble

```{r}

renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)

```


Let's look at our ts data in a couple different ways:
```{r}

renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
# renew_ts %>%  gg_season(value)

# USE THIS FOR GP GRAPHING 

ggplot(renew_parsed, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right")

```

### Just look at the hydroelectric energy consumption

```{r}

hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))

```

### What if I want quarterly average consumption for hydro?

```{r}

hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))

```

### Decompose that hydro_ts

```{r}

dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>%  autoplot()

```

```{r}

# Histogram of remainders/residuals:

hist(components(dcmp)$remainder)

```

Now let's look at the ACF:
```{r}

hydro_ts %>% ACF(value) %>% 
  autoplot()

```

### This can be dangerous:

```{r}

hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>%  autoplot(filter(hydro_ts, year(month_sep) > 2010))

```

### Make a world map

```{r}

world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"), 
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)

```

